---
title: "Update report"
format:
  html:
    toc: true
    code-fold: true
    theme: cosmo
    embed-resources: true
---

## About

This report is designed to help with providing updates to NNHIP senior team, (most likely Jenny and Katie), on changes in the most recent aggregate data submissions.

```{r}
#| label: setup
#| message: false

# sourcing

here::here("outputs", "reporting_app", "R", "dashboard_helpers.R") |> source()

here::here(".Renviron") |> readRenviron()
# readRenviron(".Renviron")

# connect to the Posit Connect board
server <- Sys.getenv("posit_server_https")
account <- Sys.getenv("posit_account")
prefix <- Sys.getenv("pin_prefix")
api_key <- Sys.getenv("posit_api_key")

board <- pins::board_connect(
  auth = "manual",
  server = server,
  key = api_key
)

pin_name <- glue::glue("{prefix}all")
pin_name_issues <- glue::glue("{prefix}issueslog")

# define a list of places where we expect to receive information
expected_places <- c(
  "Fenland and Peterborough within the North Cambridgeshire and Peterborough Care Partnership",
  "Ipswich and East Suffolk",
  "North East Essex",
  "South and West Hertfordshire (Dacorum and Hertsmere)",
  "West Essex",
  "West Suffolk",
  "Barking and Dagenham",
  "Croydon",
  "Hillingdon",
  "Kensington, Chelsea and Westminster (Bi-Borough)",
  "Lambeth and Southwark",
  "Coventry",
  "East Birmingham",
  "Herefordshire",
  "Leicestershire (West)",
  "Nottingham City",
  "Shropshire",
  "Solihull",
  "Walsall",
  "Bradford and Craven (Bradford South, Keighley and Airedale)",
  "Doncaster",
  "Leeds (Hatch, South, East)",
  "North East Lincolnshire",
  "Rotherham",
  "Stockton",
  "Sunderland",
  "Wakefield",
  "Blackburn and Darwen",
  "Morecambe Bay",
  "Rochdale",
  "Sefton",
  "St Helens",
  "Stockport",
  "Buckinghamshire (North, High Wycombe, Marlow Beaconsfield)",
  "East Berkshire and Slough",
  "East Kent",
  "East Surrey (Surrey Downs)",
  "East Sussex (Hastings and Rother)",
  "Portsmouth",
  "Bristol (South Bristol)",
  "Cornwall and The Isles Of Scilly",
  "Dorset Place ",
  "Woodspring"
) |>
  stringr::str_trim() |>
  sort()


# prepare the data
df_raw <-
  pins::pin_read(
    board = board,
    name = pin_name
  )

# process the data to make it ready for use in the app
df <-     
  df_raw |>
  # ensure there is a consistent 'metric' for each block of values
  add_metric_column_to_df() |>
  # ensure counts below threshold are suppressed
  suppress_counts() |>
  # prepare columns for use in select inputs
  factorise_columns() |>
  # indicate whether place-month has record of engagement with target cohort
  add_active_engagement_columns()

df_issues <-
  pins::pin_read(
    board = board,
    name = pin_name_issues
  )
```

```{r}
#| label: this_month

df_summary <-
  df |> 
  dplyr::distinct(place, month_zoo)

this_month <-
  df |> 
  dplyr::pull(month_zoo) |> 
  max(na.rm = TRUE)
```

# Reporting month: **`r this_month`**

## New Places submitting this month
```{r}
#| label: new_places_this_month
#| output: asis

# identify any Places submitting this month for the first time
first_time_submitters <-
  df_summary |> 
  dplyr::summarise(
    first_month = min(month_zoo),
    .by = place
  ) |> 
  dplyr::filter(first_month == this_month) |> 
  dplyr::pull(place)

# output the names of any Places submitting this month
if(length(first_time_submitters) > 0) {
  cat(paste0("- ", first_time_submitters, collapse = "\n"))
} else {
  cat("No new Places submitted this month.")
}
```

## Places missing (who've submitted before)
```{r}
#| label: old_places_missing
#| output: asis

# identify any Places who've submitted before but haven't yet this month
missing_submitters <-
  df_summary |> 
  dplyr::mutate(
    submitted_this_month = (month_zoo == this_month),
    .by = place
  ) |> 
  dplyr::summarise(
    submitted_this_month = any(submitted_this_month),
    months_submitted = paste(month_zoo, collapse = ", "),
    .by = place
  ) |> 
  dplyr::filter(!submitted_this_month) |> 
  dplyr::select(-submitted_this_month)

# display as a table
missing_submitters |> 
  gt::gt() |> 
  gt::cols_label(
    place = "Place",
    months_submitted = "Previous submissions"
  ) |> 
  gt::cols_align(
    align = "left"
  )

# display as a list (for copy-n-pasting)
if(length(missing_submitters$place) > 0) {
  cat(paste0("- ", missing_submitters$place, collapse = "\n"))
} else {
  cat("All known Places submitted this month")
}
```

## Places never submitted
```{r}
#| label: places_never_submitted
#| output: asis

# identify places which have never submitted
places_unknown <-
  setdiff(
    y = df_summary$place |> unique(), 
    x = expected_places
  )

# print as a list
if (length(places_unknown) > 0) {
  htmltools::tagList(
    htmltools::p(
      glue::glue(
        "There are {length(places_unknown)} Places which have not submitted yet:"
      )
    ),
    htmltools::tags$ul(
      lapply(places_unknown, htmltools::tags$li)
    )
  )
} else {
  htmltools::p("All Places have submitted data at least once.")
}
```

## Summary of changes / issues

### Resubmissions

```{r}
#| label: resubmissions
#| output: asis

df_resubmissions <- 
  df_issues |> 
  # get all issues logged within the last 2 weeks
  dplyr::filter(date >= (Sys.Date() - lubridate::weeks(2))) |> 
  dplyr::filter(event_type == "Resubmission")  |> 
  dplyr::mutate(month_zoo = month |> zoo::as.yearmon()) |> 
  dplyr::distinct(place_affected, month_zoo) |> 
  dplyr::arrange(month_zoo, place_affected) |> 
  dplyr::summarise(
    months = paste0(month_zoo, collapse = ", "),
    .by = place_affected
  ) |> 
  dplyr::arrange(place_affected)

if(nrow(df_resubmissions) > 0) {
  cat(paste0("- ", df_resubmissions$place_affected, ": ", df_resubmissions$months, collapse = "\n"))
} else {
  "No resubmissions received this month."
}

```


### Data caveats

This is designed to be copied into Copilot to summarise the themes:

```{r}
#| label: data_caveats
#| output: asis

df_data_caveat <-
  df_issues |>
  dplyr::mutate(month_zoo = month |> zoo::as.yearmon()) |> 
  dplyr::filter(
    month_zoo == this_month,
    event_type == "Data caveat"
  )
```


```{r}
#| label: data_caveats_places
#| output: asis
if (nrow(df_data_caveat) > 0) {
  txt <- c(
    "Places with data caveats this month:",
    "",
    paste0("- ", df_data_caveat$place_affected)
  )
  knitr::asis_output(paste(txt, collapse = "\n"))
} else {
  knitr::asis_output("No data caveats were reported this month.")
}
```

:::{.callout-tip}
The following is designed to be copied into Copilot to summarise the caveats into themes.
:::


```{r}
#| label: data_caveats_themes_prompt
#| output: asis
prompt <- c(
  "Please summarise the following data submission caveats into up to three themes.",
  "Each theme should:",
  "- capture a common underlying issue",
  "- be phrased as a short, clear heading",
  "- include a 1-2 sentence explanation",
  "",
  "Caveats:",
  paste0("- ", df_data_caveat$description)
)

knitr::asis_output(paste(prompt, collapse = "\n"))

```